Software Citation Principles

### 2016-07-26 sc-principles/rebuttal/rebuttal.tex

\item \textbf{Reviewer comment:}
\emph{I find it a bit weird to quote passages from the same paper like ``The Accessibility principle (5) states that ?....?'' (line 329 and others).}

\textbf{Our response:}
We feel it is more clear to a reader to quote the basis by which we are making follow-on comments, rather than pointing back to a previous part of the paper.

\item \textbf{Reviewer comment:}
\emph{In the discussion about Unique identification (lines 292+), I miss the discussion of hash values as version identifiers, as applied by Git for instance. As the first author is affiliated with GitHub, I assume this aspect was left out intentionally, but why? As basically all software is nowadays maintained with version control systems such as Git, these identifiers come for free and are by design unique. So, I would think they are an important aspect to discuss for software citations.}

\textbf{Our response:}
While we believe there is value in including the explicit version (e.g. Git SHA1 hash, Subversion revision number) of the software in a software citation we believe there are a number of reasons that this is not sufficient alone for the purposes of software citation:

- Version numbers are not guaranteed to be permanent: projects can be migrated to new version control systems (e.g. SVN to Git). In addition, it is possible to overwrite versions (e.g. Git force-pushing).
- Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL

``Unique identifier'' refers to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL.

We recommend that software citations capture the versions of the code expressed as (ideally) a SemVer number (http://semver.org/) as these are points in time at which the authors of the code made the explicit decision to release the code.

A SemVer release number together with a DOI (and associated archive in e.g. Zenodo) is we feel the is the most robust & future-proof solution. That said, in the absence of `recommended' identifiers such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.

We would also like to point out that there is a follow-on working group that will be producing recommended implementations for software citations in a particular setting. Ultimately the decision to (optionally) include e.g. Git SHAs are left for the implementation group to consider.

### 2016-07-29 sc-principles/rebuttal/rebuttal.tex

\textbf{Our response:}
We feel it is more clear to a reader to quote the basis by which we are making follow-on comments, rather than pointing back to a previous part of the paper.

\item \textbf{Reviewer comment:}
\emph{In the discussion about Unique identification (lines 292+), I miss the discussion of hash values as version identifiers, as applied by Git for instance. As the first author is affiliated with GitHub, I assume this aspect was left out intentionally, but why? As basically all software is nowadays maintained with version control systems such as Git, these identifiers come for free and are by design unique. So, I would think they are an important aspect to discuss for software citations.}

\textbf{Our response:}
While we believe there is value in including the explicit version (e.g. Git SHA1 hash, Subversion revision number) of the software in a software citation we believe there are a number of reasons that this is not sufficient alone for the purposes of software citation:

- Version numbers are not guaranteed to be permanent: projects can be migrated to new version control systems (e.g. SVN to Git). In addition, it is possible to overwrite versions (e.g. Git force-pushing).
- Stating a version number doesn't guarantee that the software is available at a particular (web-resolvable) URL
- When we talk about a ``Unique identifier'' in the document, we're refering to unique, persistent, and machine-actionable identifiers such as a DOI, ARK, or PURL and we don't believe a repository address with a version number is sufficient.

@@

We recommend that software citations capture the versions of the code expressed as (ideally) a SemVer number (http://semver.org/) as these are points in time at which the authors of the code made the explicit decision to release the code.

A SemVer release number together with a DOI (and associated archive in e.g. Zenodo) is we feel the is the most robust & future-proof solution. That said, in the absence of `recommended' identifiers such as a DOI) a Git SHA together with a link to a Git repository would be preferable to not having this information.

We would also like to point out that there is a follow-on working group that will be producing recommended implementations for software citations in a particular setting. Ultimately the decision to (optionally) include e.g. Git SHAs are left for the implementation group to consider.

We have updated the paper text to reflect some of these ideas.

### 2016-07-29 sc-principles/software-citation-principles.tex
is focused -> focuses
Adding discussion about versions to main text

As the process of research\footnote{We use the term ``research'' in this document to include work intended to increase human knowledge and benefit society, in science, engineering, humanities, and other areas.} has become increasingly digital, research outputs
and products have grown beyond simply papers and books to include software,
data, and other electronic components such as presentation slides, posters,
(interactive) graphs, maps, websites (e.g., blogs and forums), and multimedia
(e.g., audio and  video lectures).  Research knowledge is embedded in these
components. And papers and books themselves are also becoming increasingly
digital, allowing them to become executable and reproducible. As we move towards
this future where research is performed in and recorded as a variety of linked
digital products, the characteristics and properties that developed for books
and papers need to be applied to all digital products and possibly adjusted.
Here, we are concerned specifically with the citation of software products. The
challenge is not just the textual citation of software in a paper, but the more
general identification of software used within the research process. This work is
focused on making software a citable entity in the scholarly ecosystem. While
software products represent a small fraction of the sum total of
research output, this work together with other efforts such as the FORCE11
Data Citation Principles~\cite{data-citation-principles,10.7717/peerj-cs.1}
collectively represent an effort to better describe (and cite) all outputs of
research.

@@

>>>>

As the process of research\footnote{We use the term ``research'' in this document to include work intended to increase human knowledge and benefit society, in science, engineering, humanities, and other areas.} has become increasingly digital, research outputs
and products have grown beyond simply papers and books to include software,
data, and other electronic components such as presentation slides, posters,
(interactive) graphs, maps, websites (e.g., blogs and forums), and multimedia
(e.g., audio and  video lectures).  Research knowledge is embedded in these
components. And papers and books themselves are also becoming increasingly
digital, allowing them to become executable and reproducible. As we move towards
this future where research is performed in and recorded as a variety of linked
digital products, the characteristics and properties that developed for books
and papers need to be applied to all digital products and possibly adjusted.
Here, we are concerned specifically with the citation of software products. The
challenge is not just the textual citation of software in a paper, but the more
general identification of software used within the research process. This work
focuses on making software a citable entity in the scholarly ecosystem. While
software products represent a small fraction of the sum total of
research output, this work together with other efforts such as the FORCE11
Data Citation Principles~\cite{data-citation-principles,10.7717/peerj-cs.1}
collectively represent an effort to better describe (and cite) all outputs of
research.

@@

While we believe there is value in including the explicit version (e.g. Git SHA1 hash, Subversion revision number) of the software in any software citation we believe there are a number of reasons that a commit reference together with a repository URL is not ideal for the purposes of software citation:

\begin{enumerate}
\item Version numbers/commit references are not guaranteed to be permanent. Projects can be migrated to new version control systems (e.g. SVN to Git). In addition, it is possible to overwrite/clobber particular version (e.g. force-pushing in the case of Git).
\item A repository address and version number doesn't guarantee that the software is available at a particular (resolvable) URL.
\item A particular version number/commit reference may not represent a `preferred' point at which to cite the software from the perspective of the package authors.
\end{enumerate}
